---
title: "PS0700 - Research Exercise <br> Supplemental Guide"
# date: "`r Sys.Date()`"
date: "Created on: 2019-03-31 <br> &nbsp; &nbsp; &nbsp; &nbsp;&nbsp; Updated on: `r Sys.Date()`"

author: "Fanghui Zhao"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    css: readthedown_custom.css
---


```{r knitr_init, cache=FALSE, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

-----------------------------


# Research Assignment Guide

This is a very brief guide to help you get started on the research exercise assignment. If you are having trouble finding a suitable topic, I hope the sections on "Finding ideas" and "Formulating hypotheses" will help you to get started on the project; If you already have a topic in mind (awesome!), then you might find the latter sections more helpful. 

<div class="alert alert-info">
  This guide is subject to future updates
  
  * Check the "Updated on" date at the end of sidebar to see if there is any recent change.
  * Check the [Changelog][Changelog] section for details on the changes made.
</div>

## What you need to do (Paper outline)

This is a general paper outline you can follow to make sure you address all the required elements in the assignment.

### 1. Research topic and theory

* Focus on a general topic and  related to this general topic
  * Approach 1: Find a common $Y$, and look for two potential $X$ that explains/predicts the same $Y$
    * $X_1 \rightarrow Y$
    * $X_2 \rightarrow Y$
  * Approach 2: Find a common $X$, and look for two potential $Y$ that could be affected by this $X$
    * $X \rightarrow Y_1$
    * $X \rightarrow Y_2$
* State your theory and expectations
  * *how* do you expect the $X$ to affect $Y$, and *why*
* Provide a citation of published article/book that relates to your idea and supports your arguments on how and why your $X$ is related to $Y$.


### 2. Hypothesis and operationalization

* Formulate **two** sets of hypotheses 
* Operationalize your hypotheses using **three** variables from one of the three datasets provided (or other data source, if you prefer)
  * Be clear about the direction of the relationship
* Discuss issues of measurement validity and reliability related to the variables used in your hypotheses


### 3. Analysis and interpretation

* Of the three statistical procedures we learnt in class, choose the appropriate one to test each of your hypothesis
  * t-test
  * cross-tabulation and chi-square test
  * regression
* Use Stata to conduct the statistical test and interpret the test results
* Evaluate the substantive implications of your statistical analysis
  * Are the relationships you observe substantively/practically important?
  * What do they tell you about the way the political world works?


### 4. Discussion and future direction

* Discuss how do you think all of these variables are related in a causal sense
* Discuss potential ways to improve on your analysis in future work


## Finding ideas for a research question

If you have no ideas where to begin, try browsing through the [topics available in the three datasets](#datasets) listed below, and see if there is anything that you find interesting or stood out to you. 

Then think about whether this topic is a phenomenon that you want to explain, or do you think this subject/topic plays a role in causing/affecting other phenomenons?

I want to explain this thing!  

* Then "*this*" is your dependent variable $Y$. 
* Try to come up with two factors that could bring a change to your $Y$ -- these will be your independent variables $X$.

For example, say I want to explain voter turnout ($Y$). What could affect someone's decision to vote in an election?

* Education $\rightarrow$ Turnout 
* Interests in politics $\rightarrow$ Turnout

I think this thing explains/predicts something else! 

* Then "*this*" is your independent variable $X$. 
* Try to come up with two different phenomenons that your $X$ could bring a change to -- these will be your dependent variables ($Y$).

For example, say I think regime types (democracy or authoritarian) play an important role in shaping other outcomes. What phenomenon could be affected by regime types?

* Regime types $\rightarrow$ Income inequality
* Regime types $\rightarrow$ Gender inequality

Ta-da! Two research questions.

You might have noticed that this way of looking for research ideas is not very theoretically driven. It is fine for this exercise. However, I should emphasize that this does not mean "anything goes". Remember that you still need to provide a citation to existing work that relates to your idea. 


## Formulating hypotheses

After you have the questions you are interested in investigating, you will need to formulate two sets of testable hypotheses. 

* State *how* the two variables are related: 
  * Which variables is $X$, and which is $Y$?
* Specify what is the direction of this relationship
  * $X$ affects $Y$ positively or negatively?
* Be clear what is the unit of analysis
  * Individuals? Countries?
* Make sure the hypothesis is empirically testable
  * The null and the alternative hypothesis cannot both be true.
  
**Example 1:** 

Hypothesized relationship: Job training program $\rightarrow$ Salary change

<!-- Let $\mu$ be average salary change, $\text{treatment}$  -->

$H_0$: There is no effect of participation in job training programs on the individual’s subsequent salary

$$
H_0: \mu_{\text{treatment}} - \mu_{\text{control}} = 0
$$

$H_A$: Participation in job training programs positively increases the individual’s subsequent salary

$$
H_A: \mu_{\text{treatment}} - \mu_{\text{control}} \ne 0
$$


<!-- **Example 2:**  -->

<!-- Hypothesized relationship: Immigration status $\rightarrow$ Salary change -->

<!-- $H_0$: Immigration status and the size of salary increases are statistically independent in the population -->

<!-- $$ -->
<!-- H_0: \mu_{\text{immigrant}} - \mu_{\text{non immigrant}} = 0 -->
<!-- $$ -->

<!-- $H_A$: Immigrants are less likely to register big salary increases than non-immigrants -->

<!-- $$ -->
<!-- H_A: \mu_{\text{immigrant}} - \mu_{\text{non immigrant}} \ne 0 -->
<!-- $$ -->


**Example 2:** 

Hypothesized relationship: Years of education $\rightarrow$ Salary

$$
\text{Salary} = \alpha + \beta_0 \cdot \text{Years of education} + \epsilon
$$

$H_0$: Individuals' years of education has no effect on their current salary level

$$
H_0: \beta_0 = 0
$$

$H_A$: Individuals' years of education is positively associated with their current salary level

$$
H_A: \beta_0 \ne 0
$$


## Operationalizing variables

Find the variables in respective dataset that you want to use to measure the concepts in your hypotheses.

Make sure you know *how* the variable is measured.

* What is the unit of observation (e.g. individual or aggregate)? 
    * Be sure it corresponds to the units of analysis proposed in your hypothesis. Otherwise, you will be committing **ecological fallacy**, or the reverse **individualistic fallacy**.
* If it is measured using survey items, what is the question wording, and what are the response options?
* If using an index (created by your or already created in the dataset):
    * What individual items are used in creating the index?
    * How is the index/score created? Mean value of all items? Or sum? Or standardized score? Or something else?
* If using aggregate level data (e.g. country-level characteristics in Democracy Cross-national dataset), what is the original data source? Is it from other organizations or agencies, or from previously published papers?
    * For Norris' Democracy Cross-national dataset, see the code book (variable description or the last few pages) for the various original data sources.
    * e.g. `Export2006` is data on exports from World Bank; `frac_eth` is a measure of ethnic fractionalization based on Alesina (2003).


You will need these information 1) to decide if your variables need any further processing (e.g. recoding, creating different categories), and 2) to evaluate reliability and validity issues in your measurement.

Also pay attention to the **levels of measurement**: How your variables are measured (categorical or continuous) will determine which is the appropriate statistical procedures to use (see next section for more). 


## Conducting statistical tests

The following section provides you with a general outline of the steps involved to conduct statistical analysis, and things to consider along the way. For more concrete examples and the relevant Stata commands to use, see the sample analysis Stata do files provided during the recitation sessions.

Now, I'm going to assume that you have already decided on the variables you will be using for the analysis. 

<!-- * Data preparation -->
<!-- * Exploratory analysis -->
<!-- * Statistical analysis -->

### Preparing your data for analysis

* <span style="color:crimson">**[Must do]**</span> Look at your selected variables using commands such as `codebook`, `tab`
  * See if it is a categorical variable (i.e. nominal or ordinal), or a continuous variable (i.e. interval or ratio)
* <span style="color:lightsalmon">**[Recommend]**</span> Create informative variable names, variable labels, and value labels (for categorical variables)
* <span style="color:crimson">**[Must do]**</span> Check for missing values in your selected variables
  * Use commands such as `codebook`, `tab` to see if there are missing values
  * Cross reference respective dataset's codebook if it's not clear
  * Recode the missing values to `.` if applicable
* <span style="color:steelblue">**[If applicable]**</span> Recode the values in the variable
  * Can use `recode` to present your variable in a more intuitive way, e.g. higher value corresponds to more free, or more happy, or more supports etc.
  * Reverse code some variables if the original one is coded in a counter-intuitive way
* <span style="color:steelblue">**[If applicable]**</span> Create new variables, depending on your need
  * Create a new categorical variable from an existing continuous variable
  * Create a new categorical variable from another categorical variable
  * Create a scale/index based on multiple indicators



### Describing and exploring your data 

<span style="color:teal">**[Optional]**</span> You are not required to do the following. If you do decide to include any descriptive statistics or exploratory graphs in your paper, do not spend more than a short paragraph on it. The formal analysis part is more important.

* Check distribution shape, outliers, summary statistics
* Use summary statistics or exploeratory plots to see if there *seem* to be a relationship between your $X$ and $Y$

Depending on your variable types, you should use different strategies for exploratory analysis:

One variable distribution:

| 	| Non-graphical 	| Graphical	|
|---	|------------	|  ---------	|
| **Categorical (Nominal or Ordinal)** 	| Frequency table <br> `tab` 	| Bar plot <br> `graph bar` |
| **Continuous (Interval or Ratio)** 	| Summary statistics <br> `summarize`, `tabstat` 	| Histogram `histogram` <br> Box plot `graph box` 	|



Bi-variate relationship:

| 	| Non-graphical 	| Graphical	|
|---	|------------	|  ---------	|
| **Categorical x Categorical** 	| Frequency table <br> `tab` 	| Bar plot <br> `catplot` |
| **Categorical x Continuous** 	| Summary statistics <br> `summarize`, `tabstat` 	| Histogram `histogram` <br> Box plot `graph box` 	|
| **Continuous x Continuous** 	| Correlation <br> `corr` 	| Scatter plot <br> `scatter` 	|



### Choosing the approriate test

<span style="color:crimson">**[Must do]**</span> Depending on the types of variables used in your hypothesis, select the appropriate statistical test to determine if there is a statistically significant relationship between the $X$ and $Y$ in your hypothesis. 

| Dependent Variable \\ Independent Variable 	| IV = Nominal or Ordinal 	| IV = Interval or Ratio 	|
|-------------------------------------------	|------------------------	|--------------------------------------------------------------------	|
| **DV = Nominal or Ordinal** 	| Cross-tabulation <br> `tab` 	| Collapse continuous IV into appropriate categories, <br> then use Cross-tabulation^[Logistic regression is the proper procedure in this case, but since we did not cover it for this course, we are using this approach as a substitute.] <br> `tab` 	|
| **DV = Interval or Ratio** 	| (If IV has only two categories) T-test <br> `ttest` 	| Regression <br> `regress` 	|


Special cases when you have a continuous dependent variable, and a categorical independent variable that has more than 2 categories:

* t-test only works when independent variabls has 2 groups/categories (i.e. dichotomous)
  * more than 2 categories would involve a technique called Analysis of Variance (ANOVA), which we would not cover for this course.
* What to do then? Make a new dichotomous variable
  * Approach 1: By collapsing the exisitng categories
    * e.g. if you have a variable on individuals' highest level of educational degree (e.g. GED, high school diploma, Bachelors), you can create a College vs No college variable  by grouping degrees below college (e.g. GED, HS diploma) into one group, and degress above college into one group (e.g. Bachelors, Masters). 
  * Approach 2: By lumping different groups together into a "super-group"
    * e.g. if you have a variable on a country's geographical region that has many categories (e.g. Africa, Europe, Asia), you can create a dichotomous variable to represent (African countries vs Non-African countries) by putting all countries from "not Africa" categories into one big category.
  * Approach 3: By removing observations from certain categories
    * e.g. if the variable on voter choice has 3 categories: Clinton, Trump, Johnson/Stein, you can create a dichotomous variable (Voted for Clinton vs Voted fr Trump) by excluding the Johnson/Stein voters from analysis (recode them to missing).
* Then, use this newly create dichotomous independent variable to conduct t-test


### Interpreting test results

Important things to look out for when interpreting Stata outputs

* Direction of relationship
* Is the relationship statistically significant? What is the significance level?
* Strength of relationship
* What does the statistical results mean substantively?


<div class="alert alert-info">
To be updated after lectures on how to interpret the different test results
</div>


<!-- # Example -->



-----------------------------

# Datasets Guide: Topics and Categories {#datasets}


## American National Election Studies (ANES) 2016

To look for specific variable names as they appear in the dataset, you can either 1) open the document called ANES variable list, do a keyword search, or 2) use [one of these utility tools][Other useful tools to explore ANES].


```{r eval=FALSE, include=FALSE}
Egalitarianism

* `V162243` to `V162246`
* 4 items, take average

Racial resentment scale / Symbolic racism

* `V162211` to `V162214`
* 4 items, take average

Political efficacy

* `V162215` to `V162218`
* 4 items, take average

Moral traditionalism

* `V162207` to `V162210`
* 4 items, take average

Authoritarianism (child rearing scale)

* `V162239` to `V162242`
* 4 items, take average

<!-- TIPI (Ten Item Personality Inventory), a brief measurement of the "Big Five" personality traits -->


Stereotypes about different race and religious groups

* `V162345` to `V162356`
* There are different ways to use these items, for example:
  * Look at how individuals rate one group on one stereotype in isolation (e.g. whether they perceive this group as peaceful or violent).
  * Look at how individuals rate different groups on the same stereotype (e.g. whether they perceive one group as more violent than another group).

Sexism scale

* Modern sexism: `V162231x`, `V162232`, `V162233`
  * 3 items, take average
* Gender resentment scale (I think this set of questions is newly developed, might not have too many references on this): `V161507` to `V161510`
  * 4 items, take average


Political knowledge

* `V162072` to `V162076b`
* 5 items, take sum
* 4 of the questions are same, but are asked in different format to a random subset of the respondents (e.g. `V162073` and `V162073a`). For these 4 questions, create a new variable by combining the two different versions, then take the sum of all 5 items.

Trust in government

* `V161215` to `V161218`
* 4 items, take average
```

### Partisanship and attitudes towards parties

* Feelings about parties
* Party identification
* Closeness to parties
* Party performance
* Role of parties

### Attitudes towards candidate and incumbent

Affective attitudes and evaluations

* Feeling thermometers
* Candidate evaluations
* Traits of president, candidates
* Affect toward President, candidates
* Evaluation of Congressional candidates

Performance evaluations

* Performance of current president
* Retrospective evaluations
* Other government performance
* Candidate performance
* Congressional candidate performance
 

### Issues

* General issues (most important problems, country on the right track)
* Social welfare issues (e.g. healthcare, jobs, income)
* Racial policy issues
* Economic issues
* Foreign relations
* Social issues (e.g. abortion, gun control, immigration)
* Environment
* Federal spending 

### Ideology and values

* Liberal/Conservative
* Religious values
* Moral traditionalism `V162207` to `V162210`
* Egalitarianism `V162243` to `V162246`
* Racial resentment `V162211` to `V162214`
* Social trust 
* Authoritarianism (child rearing scale) `V162239` to `V162242`
* Modern sexism: `V162231x`, `V162232`, `V162233`
 
### System support

General attitudes toward the political system, the nation, or the government.

* Trust in government `V161215` to `V161218`
* Power of the federal government
* Political efficacy `V162215` to `V162218`
* Government responsiveness
* Patriotism
* Feeling thermometers to different branches of government


### Political participation and mobilization

Civic participation

* Involvement in community
* Attitudes towards voting (duty or choice)

Engagement and information

  * Interest
  * Discuss politics
  * Political knowledge `V162072` to `V162076b`

Campaign activity

  * Respondent’s campaign activity
  * Campaign contact/solicitation

Registration, turnout, vote choice

  * Previous Presidential election
  * Vote intention (pre-election)
  * Turnout & registration
  * Vote choice
  * State primary/caucus 
 
### Media

Use of news media, such as television, the Internet, newspapers, and radio. 

### Social groups

* Feeling thermometers to various groups
* Group identification and closeness
* Group stereotypes `V162345` to `V162356`


### Socio-demographic data

* Education
* Employment status
* Religious identification
* Race/ethnicity
* Age
* Mobility
* Income
* Social class
* Other personal and demographic data (labor union, home tenure, marital status, etc.) 


### Other useful tools to explore ANES

Useful tools to explore variables in ANES dataset based on categories

* [SDA utility from UC Berkeley](http://sda.berkeley.edu/sdaweb/analysis/?dataset=nes2016)
  * Navigate the categories to explore which variables are available, and which ones are measuring similar concepts. Variable name for each question is also given, so you can find the corresponding variable in the dataset. This is useful when you are not very familiar with the ANES data, and not sure exactly what data are available.
* [ANES continuity guide](http://electionstudies.org/resources/anes-continuity-guide/)
  * Note: not all questions are available in the 2016 dataset, check the key. If the key is shaded for year 2016, then it is available. Mouse over to see the variable name, and use that to locate the variable you want in the dataset.


-----------------------------


## Democracy Cross-national Dataset (Pippa Norris)

The following categories are taken from the table of content on Norris' code book. I rearranged the orders to group the similar categories together, for ease of search. To find the variables available in each category, open the code book and do a keyword search for the category name, e.g. "Decentralization", or refer to the table of content to look for the corresponding page numbers ("Decentralization" on p.23).

Measures of democracy:

* Standardized Democracy Indices
* Types of Regimes by Freedom House
* Democracy and Dictatorship Classification
* Polity Classifications

Institution characteristics:

* Typology of Executives
* Decentralization
* Corruption Perception Index
* Worldwide Governance Indicators (World Bank)
* Classifications of Ethnic Pluralism

About elections:

* Classification of Electoral System
* Gender Political Representation
* Turnout
* Electoral Integrity
* IDEA Political Finance Database
* Database of Political Institutions (World Bank)


Human rights and development indicators:

* Human Rights
* UNDP Human Development Indicators
* World Bank Development Indicators

Media 

* Media Freedom
* Media Access
* Media ownership
* Cultural export

Social, economic, environmental variables:

* Globalization Index and Cultural Trade
* Aid
* Religion Variables
* World Social, Economic and Environmental Variables from Earlier Years

Aggregate survey data

* World Values Survey (aggregated value by nation)

-----------------------------

## World Value Survey (WVS) Wave 6 (2010-2012)

Here is a selected subset of WVS variables that (I think) are more directly related to politics. There are many more variables in the WVS. Feel free to explore more. 

I've indicated some of the variable names as they appear in the dataset, and many other I have not. You can find the corresponding variable name by search for the category name in the WVS code book (not the questionnaire!).


### Politics

* Political action
* Political system (which type is good/bad)
* Aims of country (what goals are most important)
* Interest in politics
* Protecting environment vs. Economic growth
* Idelogy (left/right) `V95`
* Views on various issues `V96` to `V100`
* Democracy (what are the essential characteristics) `V131` to `V139`
* Information source
* Vote in elections
* Perceptions of electoral integrity`V228A` to `V228I`
* National identity `V211` to `V216`

### Society

* Confidence (in various institutions) `V108` to `V126`
* Membership in various groups and organizations `V25` to `V35`
* Tolerance towards minorities (groups of people you would not like to be neighbors with) `V36` to `V44`
* Trust: `V102` to `V107`
* Security in neighborhood `V170` to `V178`
* Views on science `V192` to `V197`


### Family and Work

* When jobs are scarce `V45`, `V46`
* Views on gender role/women and work `V47` to `V54`


### Religions and morals

* Various questions on religion `V144` to `V156`
* Justifiable (are various kinds of actions justifiable) `V198` to `V210`


### Special indexes

Materialist/Post-materialist index (a.k.a. Inglehart Index, based on research by Ronald Inglehart).

* `Y001`: 12 item summary index, created based on questions `V60` to `V65`
* `Y002`: 4 item summary index, created based on questions `V62` and `V63`

Autonomy index `Y003`

* Important qualities children should learn
* Created from `V12` to `V22`



-----------------------------

# Changelog

2019-04-05

* Edited "What you need to do" section
* Expanded on "Conducting statistical tests" section

2019-04-01

* Initial release.